### 2017-05-08 paper.tex
Set up paper, added intro and part of technical description

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\graphicspath{{./figs/} {./plots/}}

\usepackage{balance}
\usepackage{comment}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage{float}
\newcommand{\sys}{Safewalk\xspace}
\author{
  \IEEEauthorblockN{
   Iljoo Baek, Mengwen He and Emily Ruppel
  }
 \IEEEauthorblockA{
   Carnegie Mellon University
  }
 \IEEEauthorblockA{
   {ibaek,mengwenhe,eruppel}@andrew.cmu.edu
  }
}
\title{Gait Monitoring using Wireless Sensors\\
        \vspace{4mm}
        \large{Final Report 18-748}\\
        \large{Spring 2017}}

\begin{document}

\maketitle

%Only including an abstract since it looks like the submission site wants one... 
\begin{abstract}
Killer abstract. 
\end{abstract}

\section{Introduction}
  \label{sec:intro}
  Gait patterns, the measurable characteristics of an individual’s walking movement, can
provide insight into a variety of components of an individual’s health status. Certain
gait pathologies may be associated with specific medical conditions. Sudden changes in the
way an individual walks can be indicative of the occurrence of an undetected but
deleterious medical event, i.e. undetected stroke. A shift in gait pattern over time may
come about as the result of muscular or neural decline that could increase the probability
of a dangerous event, such as a fall, in the future. However, diagnosing conditions or
predicting an increased fall risk from gait patterns requires the attention of a trained
medical professional, which limits the settings in which gait analysis can be used.

Wireless Body Area Networks have been proposed as a promising solution for capturing
useful health status indicators without inconveniencing a patient~\cite{wban}. The
additional data enables a new field of computer-assisted physical therapy that allows
therapists to deliver more targeted care to their patients. Advanced treatment centers for
patients with unique cases, such as the Walter Reed Center for Performance and Clinical
Research (CPCR) amputee clinic, use extensive computer modeling to aid in patient
rehabilitation~\cite{cpcr}.  An apparent barrier to the widespread adoption of computer
assisted therapy is the cost of the motional analysis systems. For instance, the CPCR uses
23 infrared cameras to track the movement of reflective markers attached to a patient's
body and six force plates in the floor to fully capture a patient's movement. Though the
movement data can be captured and used at a later time, the gait analysis is still limited
to a clinical setting.

Our project's goal was to expand the usefulness of gait analysis beyond a professional’s
office to an individual’s day-to-day environment. We accomplished our goal by
demonstrating a proof-of-concept gait monitoring system that relies on wireless motes
instrumented with Inertial Measurement Units (IMUs) to capture relevant data and relay
them to a powerful backend machine for offline and real time processing. Our system
extracted a variety of components of the subject's gait pattern from linear and angular
acceleration data measured at specfic points along his or her legs. The subject's
movements were then visualized and features of the gait were plotted in real time. We also
demonstrate the practicality of real-time fall detection by using a simple thresholding
technique to indicate when a subject is no longer upright. Throughout the trials with our
prototype system we note that wireless sensor networks are particularly promising for
accomplishing gait detection because the small nodes and lack of constraining wires
minimizes the impact on the patient and allows for an honest gait assessment without the
residual impact of cumbersome measurement devices. 
%
%Add note about final system architecture in Intro... should make the technical
%description easier and it still make sense to do that here
%
\section{Related Work}
This project is based on prior work in gait assessment, IMU data interpretation, and
wireless communication. In this section we will describe how the prior work influenced the
design of our project

\section{Technical Description}
To implement our gait monitoring system we had to handle three major design points-
transmitting data from the IMUs, analyzing the IMU data, and building a useful interface
to the information. This section will describe the choices our team made with regards to
these three subsystems. 
\subsection{Device Communication}
Moving data from the IMU to the remote server can be separated into two distinct
transitions-- one from IMU to firefly slave device and another from the slave to the
master node. As shown in Figure~\ref{fig:ff}, the IMU board attaches directly to the
Firefly. The connection provides power from the Firefly batteries to the IMU and allows
for two way communication over one of the Firefly's auxiliary UART ports and the
configurable IMU communication ports. The majority of the data transported over the serial
connection is the IMU readings transported from the IMU board to the Firefly whenever the
IMU has data available.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cropped2}
  \caption{{\bf Firefly node with IMU board attached}}
  \label{fig:ff}
\end{figure}

  The format of the packet passed from IMU to Firefly is shown in
Figure ~\ref{fig:packet}. The data are encoded as ASCII characters to simplifiy debugging
as data move from IMU to firefly slave node to master and final output. While the ASCII
encoding greatly simplified debugging, it does mean that the packet size may change given
different accelerometer values.  Throughout the project, several unsuccessful attempts
were made to tightly control the data transmission by the IMU board, but the nature of the
UART protocol made timing discrepancies between the two devices difficult to overcome, and
was further complicated by the variable packet size. Instead, we settled on a four
character pattern to indicate the beginning of an IMU transmission followed by a single
special character to indicate the end. In contrast to the long packets transmitted from
the IMU board, the firefly needs only to issue commands to the IMU, so the protocol from
firefly to IMU is far simpler. If the IMU board receives a character from the Firefly, the
corresponding command is retrieved from a look-up table and carried out. A final feature
of the IMU-Firefly communication protocol is automatic reboot of the firefly. 
  Once packets are passed from the IMU to the Firefly slave device, the firefly transmits
the packet using the Point Coordination Function (PCF) TDMA protocol~\cite{pcf}. Each
slave device is statically assigned a time slot according to its preprogrammed MAC address
and only transmits during its assigned slot. The PCF TDMA protocol is designed for a
network with a star topology, where all slave nodes are within one hop of the master
device because all slave nodes must synchronize based on a message sent from the master at
the beginning of every TDMA cycle.  Despite its simplicity, this protocol met our design
goals of maximizing data throughput by preventing collisions without inucurring the time
overhead of carrier sense~\cite{CSMA}.  During its designated time slot, a slave node
transmits the packet passed to it from the IMU after stripping off the start and stop
bytes. An added advantage of the TDMA protocol is the implicit ID and time information
coupled with each transmission. Based on the time slot during which a packet is received,
the master node can automatically attach the ID of the node that transmitted the packet
before passing the entire packet to the server for processing. Further, the time
synchronization as a result of TDMA allows for the assumption that the measurements from
different IMUs received during a given TDMA cycle occurred at the same time. Therefore,
the server can afix a timestamp to data based on when it receives the data rather than
transmitting a timestamp from the slave node indicating the precise moment when the IMU
measurement was made.

The protocol delivers a data rate of approximately 10 packets per second from each of
the slave nodes. This rate could be improved by decreasing the packet size, and more
tightly matching the TDMA slot length to the transmission time of a packet. In the current
configuration, the slot length is fixed at 10ms per node. Further exploration of the
transmit and receive task lengths on the slave nodes could also improve throughput.
Currently both the receive and transmit tasks operate on a period of 250ms. In the
transmit task, a slave device transmits as many times as possible given the tdma protocol
before switching to the receive task, but during the receive task the node will receive a
message with low probability because the master sends commands to the slaves infrequently.
Reducing the length of the receive period will thererfore reduce the slave node's idle
time with little impact on the system's performance.

  The only command transmitted from the master to the slaves beyond the TDMA synchronization
is a "Recalibrate" command that signals the slave nodes to send messages to their IMUs to
perform static calibration. This command is sent only when a users pushes the auxiliary
button the master firefly at the beginning of a long period of system operation. While the
communication protocol is complicated by this single command, it is important that the
master node be able to issue a synchronous recalibration command to all of the nodes to
improve the accuracy of the following measurements. The recalibraton phase samples the
baseline static noise floor of each IMU, so the subject wearing the devices must be
stationary while the devices are undergoing recalibration. It is difficult for the subject
to remain stationary if each of the devices must be manually turned off and then on to initiate
recalibration. The master's recalibrate command circumvents the issue by causing all IMUs
to recalibrate instantaneously which greatly reduces the amount of time the patient must
remain still.

\subsection{Gait Analysis}

\subsection{User Interface}

\section{Results}
Add picture of final enclosure. Discuss how it was attached to the subject. Mention that
it was fairly stable.

Talk about all of the calculations and provide examples from data we have lying around.

\section{Conclusion}

\balance
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
\end{document}

### 2017-05-09 paper.tex
Added abstract and bundle of figures

@@
 \begin{abstract}
-Killer abstract. 
+In this project we demonstrate a gait monitoring system comprised of Firefly wireless
+sensor motes that relay data from inertial measurement units mounted on a subject's body
+to a central server. Using the acceleration and orientation information reported at six
+different points along a subject's legs and feet, we were able to visualize the subject's
+movements in real time and capture the relevant motion data to be used for future
+analysis. Our system also provides a low-processing solution for fall detection through
+thresholding analysis of the orientation measurements reported by a subset of the sensor
+nodes in the system. After testing our system on three different subjects, we believe it
+represents a proof of concept of low cost gait monitoring systems that will improve the
+quality of patient care in a multitude of settings. 
 \end{abstract}
@@
 \section{Related Work}
-This project is based on prior work in gait assessment, IMU data interpretation, and
-wireless communication. In this section we will describe how the prior work influenced the
-design of our project
+This project is based on prior work in gait analysis and wireless sensors networks used to
+aid healthcare objectives. In this section we will describe how the prior work influenced
+the design of our project 
+\subsection{Gait Analysis}
+Gait analysis using wearable sensors is a well-studied approach to characterizing gait
+patters~\cite{gaits}. The goal of the gait analysis approaches is to analyze the various phases of
+the human gait to identify abnormalities. A normal walking gait is comprised of eight
+stages as shown in Figure~\ref{fig:phases}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{figs/phases}
+  \caption{{\bf Phases of the stance (a) and swing (b) periods~\cite{gaits} }}
+  \label{fig:phases}
+\end{figure}
+
+Since the first study using wearable sensors in 1973, researchers have captured the data
+from these phases using a variety of transducers. Recently, multiple studies have used
+IMUs attached to a subject’s thigh, calf, and foot to study gait kinematics. Gait
+kinematics investigates the movement of segments of the lower limbs while walking or
+running~\cite{run}. This is distinct from gait kinetics which analyzes the forces exerted on
+different joints while walking. The difficulty with kinetic analysis is the hardware
+required to capture accurate measurements. Previous studies have relied on specialized,
+instrumented footwear that measures the forces acting on a patient’s foot. Gait
+kinematics, on the other hand, has moved towards less intrusive measurement techniques.
+For instance, ~\cite{walkers} used small wireless sensor nodes to report information about the gait
+characteristics of test subjects. Graurock, et. al went one step further in targeting
+patient convenience by building a "plug-and-play" system wherein each sensor node self
+calculated its position on a patient's lower extremeties~\cite{pairing}.
 
 \section{Technical Description}
-To implement our gait monitoring system we had to handle three major design points-
-transmitting data from the IMUs, analyzing the IMU data, and building a useful interface
-to the information. This section will describe the choices our team made with regards to
-these three subsystems. 
+The system architecture we employ to achieve our goal of measuring a subject's gait using
+wireless sensors is shown in Figure~\ref{overview}. Note that it contains six Firefly
+slave devices each with a Sparkfun Razor 9 DoF Razor IMU M0 board attached. The data are
+then transmitted from the slave nodes to a master node which finally pushes the data to a
+server. To implement our gait monitoring system we had to handle three major design
+points- transmitting data from the IMUs, analyzing the IMU data, and building a useful
+interface to the information. This section will describe the choices our team made with
+regards to these three subsystems.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.75\columnwidth]{figs/sysarch}
+  \caption{{\bf High level system design}}
+  \label{fig:overview}
+\end{figure}
+
 \subsection{Device Communication}
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{figs/cropped2}
+  \includegraphics[width=0.75\columnwidth]{figs/cropped2}
   \caption{{\bf Firefly node with IMU board attached}}
@@
 of the IMU-Firefly communication protocol is automatic reboot of the firefly.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{figs/packet}
+  \caption{{\bf Packet Format from IMU to Firefly}}
+  \label{fig:packet}
+\end{figure}
+
 Once packets are passed from the IMU to the Firefly slave device, the firefly transmits
@@
 \subsection{Gait Analysis}
+  Paragraph about step detection
+
+  Paragraph about stride length based on step detection and assuming the subject's limb
+  measurements are known.
+
+  Paragraph about calculating knee extension
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{figs/knees}
+  \caption{{\bf Knee extension Calculation}}
+  \label{fig:knees}
+\end{figure}
 

### 2017-05-10 paper.tex
Added figures, more related work, more technical description

@@
 \usepackage{float}
+\usepackage{subfigure}
 \newcommand{\sys}{Safewalk\xspace}
@@
 
+\subsection{IMU Data Processing}
+A large amount of work has been performed in processing the acceleration and orientation
+values reported by inertial measurement sensors to make use of the data. Specifically in
+terms of gait analysis, the work comes from both the robotics and motion capture
+communities. 
+%Can we get a citation from the robotics community here too? 
+For instance, Yost labs has characterized the different joint types of the
+human body in terms of the necessary interpetation of data from IMUs mounted on adjacent
+limbs~\cite{yost}. The complexity of the analysis ranges from the simple calculation to
+determine the angle of a hinge joint to the vector transformations required to assess the
+pitch, yaw and roll of multi-axis joints. Complementary work from the medical community
+puts the raw measurements in context for clinicians by providing a point of comparison.
+For instance, in ~\cite{pop} inertial sensor data was collected from a large population
+of adults and translated into a database of typical gait parameter values that researchers
+and healthcare professionals can use as a point of comparison. The study combines a
+multitude of well characterized algorithms for assessing gait parameters using only two
+IMUs, one on each of the subject's feet, and is able to build a model for normal gaits
+in terms of parameters such as foot clearance and step cadence.
+
 \section{Technical Description}
-The system architecture we employ to achieve our goal of measuring a subject's gait using
-wireless sensors is shown in Figure~\ref{overview}. Note that it contains six Firefly
-slave devices each with a Sparkfun Razor 9 DoF Razor IMU M0 board attached. The data are
-then transmitted from the slave nodes to a master node which finally pushes the data to a
+Our system follows in the spirit of prior work using IMUs to measure gait parameters, and
+work that uses wireless sensors to improve the quality of patient care. The system
+architecture we employ to achieve our goal of measuring a subject's gait using wireless
+sensors is shown in Figure~\ref{overview}. Note that it contains six Firefly slave devices
+each with a Sparkfun Razor 9 DoF Razor IMU M0 board attached. The data are then
+transmitted from the slave nodes to a master node which finally pushes the data to a
 server. To implement our gait monitoring system we had to handle three major design
@@
 \subsection{Gait Analysis}
-  Paragraph about step detection
-
-  Paragraph about stride length based on step detection and assuming the subject's limb
-  measurements are known.
-
-  Paragraph about calculating knee extension
+Our work analyzing subject's gait parameters focused on stride length, swing time, and
+knee extension. Each of these parameters required different analysis techniques that will
+be described below.\\ 
+{\bf Stride Length} To determine the stride length, we first assume that the lengths of each
+segment of a subject's limbs are known. Then...\\
+%Need to know how this works!!
+{\bf Swing Time} Our analysis focuses on detecting the subject's step to in turn ascertain
+when a subject's foot is on the ground, and when it is in motion. By making the assumption
+that the acceleration of the foot is close to zero when it is on the ground, we were able
+to detect the beginning and end of a subject's stride based strictly on acceleration data
+from the x-y plane. Our implementation is a simplified version of the technique used
+in~\cite{fancy}. Figure~\ref{fig:step-detect} shows the basic shape of accelerometer data
+gathered from a subject's feet as he or she walks and the marked regions indicate the time
+during which his/her foot is either on the ground or in the air. 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{figs/swing}
+  \caption{{\bf Step detection from accelerometer data}}
+  \label{fig:step-detect}
+\end{figure}
+\\
+{\bf Knee Extension} A subject's knee extension was calculated using the hinge joint
+techniques explained by Yost~\cite{yost}. The equations are: 
+%Insert equations here!!
 
@@
 
+\subsection{Enclosure} 
+To ensure that the IMU remains in a fixed position on the body, and to protect the
+connection between the Firefly and the IMU, we designed a 3D printed enclosure for the
+slave nodes. Figure~\ref{fig:box} shows the enclosure with and without the specially designed
+lid. The enclosure allows for easy access to all programming ports, power switches and
+batteries without time consuming disassembly. Further, the design allows a strap to easily
+pass through the upper and lower pieces of the enclosure so that the devices can be
+comfortably affixed to subjects using velcro straps. During testing, we found that the
+straps held the slave devices securely in place and they did not restrict the subject's
+movements. 
+%Except for that part where the ones on the thighs acted like tourniquets... but we won't
+%talk about that... 
+\begin{figure}[h]
+  \centering
+  \subfigure[]{
+    \centering
+    \includegraphics[width=.45\columnwidth]{figs/box}
+    \label{fig:box2}
+  }
+  \subfigure[]{
+    \centering
+    \includegraphics[width=.45\columnwidth]{figs/wLid}
+    \label{fig:wLid}
+  }
+  \label{fig:box}
+  \caption{{\bf Enclosure for sensor nodes}}
+\end{figure}
+
 \section{Results}
-Add picture of final enclosure. Discuss how it was attached to the subject. Mention that
-it was fairly stable.
 

### 2017-05-10 paper.tex
Added user interface description and utodo's

@@
+\newcount\draft\draft=1 % set to 0 for submission or publication
+\newcount\cameraready\cameraready=1
+
 \documentclass[conference]{IEEEtran}
@@
 \usepackage{xspace}
+\usepackage{xxx}
 \usepackage[export]{adjustbox}
@@
 \newcommand{\sys}{Safewalk\xspace}
+\newcommand{\utodo}[2]{ {\textcolor{red} {#1}}: {\textcolor{red} {#2}}}
 \author{
@@
   \centering
-  \includegraphics[width=0.75\columnwidth]{figs/cropped2}
+  \includegraphics[width=0.6\columnwidth]{figs/cropped2}
   \caption{{\bf Firefly node with IMU board attached}}
@@
 {\bf Stride Length} To determine the stride length, we first assume that the lengths of each
-segment of a subject's limbs are known. Then...\\
-%Need to know how this works!!
+segment of a subject's limbs are known. \utodo{Alex}{Describe how stride length was
+calculated}
+\\
 {\bf Swing Time} Our analysis focuses on detecting the subject's step to in turn ascertain
@@
 gathered from a subject's feet as he or she walks and the marked regions indicate the time
-during which his/her foot is either on the ground or in the air. 
+during which his/her foot is either on the ground or in the air. \\
 \begin{figure}[h]
@@
 \end{figure}
-\\
 {\bf Knee Extension} A subject's knee extension was calculated using the hinge joint
-techniques explained by Yost~\cite{yost}. The equations are: 
-%Insert equations here!!
+techniques explained by Yost~\cite{yost}. The equations are \utodo{Iljoo}{Add the
+equations you actually used for knee extension} Figure~\ref{fig:knees} shows the angles
+that are reported by the final calculation.
 
@@
 \subsection{User Interface}
+\utodo{Alex}{Add all the details about the user interface that I missed!!}
+The user interface is aimed at a technical professional rather than a healthcare provider,
+but it is built so that certain features are easy to access for non-technical personnel
+and data storage for future processing is streamlined. Our design uses the popular Robot
+Operating System (ROS) platform to control data flow into the system software from the
+single stream of data coming from the master node. The control buttons for data from the
+individual nodes are simple to access and can be used to change the system's
+configuration. The interface also contains a slave node health status indicator that
+reports whether or not the master node has successfully received data from a given slave
+within a recent timeframe. Figure~\ref{fig:ROS} shows the configuration interface and the
+health status indicator. In a healthcare environment, these tools would be used by
+technical personnell to configure the system software before it is used by a healthcare
+provider.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{figs/gui}
+  \caption{{\bf ROS user interface}}
+  \label{fig:gui}
+\end{figure}
+The user interface also provides a convenient video playback feature than can replay all
+of the data visualizations produced in real time during active data capture as well as an
+accompanying video of the subject. Finally, the data are logged automatically for future
+use or offline analyses.
 
@@
 \section{Results}
-
-Talk about all of the calculations and provide examples from data we have lying around.
+\
 

### 2017-05-10 paper.tex
Added results section and a bundle of figures

@@
 \section{Results}
-\
+Our results are best demonstrated through a series of the screenshots acquired during
+multiple demos. They capture the objective of our project and demonstrate the value of the
+current prototype as well as the points where improvements could be made. 
+\subsection{Visualization}
+Our final data visualization included a simple avatar and representations of the
+orientation of all six slave nodes. Figure~\ref{fig:vis} shows the final visualization layout.
+Both elements of the visualization moved in real time as data passed in from the master.
+For the orientations, each node's axes move independent of the other nodes, so the
+movements provide the user with a sense of the raw noise included in the orientation
+measurements. If any of the node's IMUs reports some kind of spurious, outlying data
+point, the user will observe a sudden change in that IMU's axis representation. Future
+work would filter out sudden changes to prevent this phenomena.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\columnwidth]{figs/vis}
+  \caption{{\bf Final visualization layout}}
+  \label{fig:vis }
+\end{figure}
+
+\subsection{Gait parameter graphs}
+In adition to visualizing a subject's movements, our system can generate graphs of a
+subject's knee extension and stride length in real time. Figure~\ref{fig:extend} shows the
+graph of knee extension on the right as our subject bends his knee. The graph contains
+separate extension angles for the left and right knee, so it is simple to notice if the
+range of motion in one knee is more restricted than the other.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\columnwidth]{figs/extend}
+  \caption{{\bf Real time knee extension reporting}}
+  \label{fig:extend }
+\end{figure}
+
+The stride length graph appears in nearly real time, but requires additional processing
+of the segments of multiple data points to find where steps occur, thus delaying it
+slightly behind real time. However, it still provides useful, prompt feedback on the
+length of a subject's stride as shown in Figure~\ref{fig:length}. The graph shows the
+distance between the right and left foot, so the rise and fall pattern observed in the
+graph in this figure indicates the feet separating as a subject's leg moves forward and coming
+back together as the subject swings one foot past the other on the next step.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\columnwidth]{figs/length}
+  \caption{{\bf Stride length reporting}}
+  \label{fig:length}
+\end{figure}
+
+\subsection{Fall Detection}
+In testing our gait monitoring system, we found that the orientation reported by the nodes
+strapped to a patient's thighs provided a reliable fall indicator for our subject. When
+the subject is walking, the measurement is stable and indicates that the node is
+perpendicular to the ground plane. If a subject falls, the orientation data indicates a
+rapid change in the angle between the node's central axis and the ground as demonstrated
+in Figure~\ref{fig:falling}. Though our system relies primarily on the obvious change in
+the grph that a system operator can visually observe, an automatic mechanism based on a
+set threshold for the angle between the nodes and the ground would suffice to warn
+caregivers that a patient has fallen.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.99\columnwidth]{figs/falling}
+  \caption{{\bf Angle reporting for fall detection }}
+  \label{fig:falling}
+\end{figure}
 

### 2017-05-10 paper.tex
Updated bibliography and makefile

@@
 \maketitle
+\thispagestyle{plain}
+\pagestyle{plain}
 
@@
 \begin{abstract}
-In this project we demonstrate a gait monitoring system comprised of Firefly wireless
+In this project we demonstrate \sys, a gait monitoring system comprised of Firefly wireless
 sensor motes that relay data from inertial measurement units mounted on a subject's body
@@
 Wireless Body Area Networks have been proposed as a promising solution for capturing
-useful health status indicators without inconveniencing a patient~\cite{wban}. The
+useful health status indicators without inconveniencing a patient~\cite{msban,wban}. The
 additional data enables a new field of computer-assisted physical therapy that allows
@@
 
-Our project's goal was to expand the usefulness of gait analysis beyond a professional’s
+\sys's goal was to expand the usefulness of gait analysis beyond a professional’s
 office to an individual’s day-to-day environment. We accomplished our goal by
@@
 residual impact of cumbersome measurement devices. 
-%
-%Add note about final system architecture in Intro... should make the technical
-%description easier and it still make sense to do that here
-%
 \section{Related Work}

### 2017-05-10 paper.tex
Added conclusion and milestones

@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{figs/gui}
+  \includegraphics[width=\columnwidth]{figs/gui}
   \caption{{\bf ROS user interface}}
@@
 \section{Conclusion}
+\sys is a proof-of-concept implementation of gait monitoring using low cost IMUs and
+a small number of wireless sensors. We were able to measure relevant parameters of our
+subject's gaits and visualize the data in an accessible fashion despite the relatively low
+data transfer rate in our network. \sys represents an example of a low cost monitoring
+solution for patients with fall risks, and it could provide additional data for healthcare
+providers seeking to customize a plan of care to a patient's behavior outside of the
+office. We worked through several major challenges including timing mismatch between the
+IMU and the Firefly, packet loss in the network, IMU drift, and signal processing
+subtleties.
+\section{Milestones and Task Division}
+We achieved most of the milestones that we set out to accomplish in our original project
+proposal, though some of the deadlines for the different milestones shifted.
+Figure~\ref{fig:table} is a timeline of our original deliverable schedule and an explanation
+of when each deliverable was finished or how it changed.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{figs/table}
+  \caption{{\bf Timeline of deliverables }}
+  \label{fig:table}
+\end{figure}
+
+The table below reports the segments of the project each team member completed.\\
+\begin{tabular}{|p{.95\columnwidth}|}
+  \hline
+{\bf Alex:}
+\begin{itemize}
+  \item {IMU calibration feasibility assessment}
+  \item ROS system development and management
+  \item Motion capture and visualization
+\end{itemize}\tabularnewline
+\hline
+{\bf Emily:}
+\begin{itemize}
+  \item{Data transfer from IMU to Firefly}
+  \item{Inter-Firefly communication management}
+  \item{Final report organization} 
+\end{itemize}\tabularnewline
+\hline
+{\bf Iljoo:}
+\begin{itemize}
+  \item{Knee extension calculation} 
+  \item{Step detection research and implementation}
+  \item{Enclosure fabrication} 
+\end{itemize}\tabularnewline
+  \hline
+\end{tabular}
 

### 2017-05-10 paper.tex
Added to technical description, ran spellcheck, fixed typos

@@
 patient convenience by building a "plug-and-play" system wherein each sensor node self
-calculated its position on a patient's lower extremeties~\cite{pairing}.
+calculated its position on a patient's lower extremities~\cite{pairing}.
 
@@
 For instance, Yost labs has characterized the different joint types of the
-human body in terms of the necessary interpetation of data from IMUs mounted on adjacent
+human body in terms of the necessary interpretation of data from IMUs mounted on adjacent
 limbs~\cite{yost}. The complexity of the analysis ranges from the simple calculation to
@@
 architecture we employ to achieve our goal of measuring a subject's gait using wireless
-sensors is shown in Figure~\ref{overview}. Note that it contains six Firefly slave devices
+sensors is shown in Figure~\ref{fig:overview}. Note that it contains six Firefly slave devices
 each with a Sparkfun Razor 9 DoF Razor IMU M0 board attached. The data are then
@@
   The format of the packet passed from IMU to Firefly is shown in
-Figure ~\ref{fig:packet}. The data are encoded as ASCII characters to simplifiy debugging
-as data move from IMU to firefly slave node to master and final output. While the ASCII
+Figure ~\ref{fig:packet}. The data are encoded as ASCII characters to simplify debugging
+as data move across the multiple platforms in our system. While the ASCII
 encoding greatly simplified debugging, it does mean that the packet size may change given
@@
 the beginning of every TDMA cycle.  Despite its simplicity, this protocol met our design
-goals of maximizing data throughput by preventing collisions without inucurring the time
+goals of maximizing data throughput by preventing collisions without incurring the time
 overhead of carrier sense~\cite{CSMA}.  During its designated time slot, a slave node
@@
 different IMUs received during a given TDMA cycle occurred at the same time. Therefore,
-the server can afix a timestamp to data based on when it receives the data rather than
+the server can affix a timestamp to data based on when it receives the data rather than
 transmitting a timestamp from the slave node indicating the precise moment when the IMU
@@
 Currently both the receive and transmit tasks operate on a period of 250ms. In the
-transmit task, a slave device transmits as many times as possible given the tdma protocol
+transmit task, a slave device transmits as many times as possible given the TDMA protocol
 before switching to the receive task, but during the receive task the node will receive a
 message with low probability because the master sends commands to the slaves infrequently.
-Reducing the length of the receive period will thererfore reduce the slave node's idle
+Reducing the length of the receive period will therefore reduce the slave node's idle
 time with little impact on the system's performance.
@@
 master node be able to issue a synchronous recalibration command to all of the nodes to
-improve the accuracy of the following measurements. The recalibraton phase samples the
+improve the accuracy of the following measurements. The recalibration phase samples the
 baseline static noise floor of each IMU, so the subject wearing the devices must be
@@
 be described below.\\ 
-{\bf Stride Length} To determine the stride length, we first assume that the lengths of each
-segment of a subject's limbs are known. \utodo{Alex}{Describe how stride length was
-calculated}
-\\
+{\bf Stride Length} To determine the stride length, we first assume that the lengths of
+each segment of a subject's limbs are known. Then we use the attached IMUs to measure
+their orientations. After getting the length and direction of each limb, we can simulate
+the gait motion using a model of linked limbs with a fixed rotation axis of the hip. To do
+that, we can first calculate the 3D position of the knee using the measurement of the
+thigh, and then calculate the 3D position of ankle using the measurement of the calf and
+the estimated knee position. Furthermore, if we can detect the step, we can calculate the
+stride length equals the longitudinal distance between two ankle points.\\
 {\bf Swing Time} Our analysis focuses on detecting the subject's step to in turn ascertain
@@
 to detect the beginning and end of a subject's stride based strictly on acceleration data
-from the x-y plane. Our implementation is a simplified version of the technique used
-in~\cite{fancy}. Figure~\ref{fig:step-detect} shows the basic shape of accelerometer data
-gathered from a subject's feet as he or she walks and the marked regions indicate the time
-during which his/her foot is either on the ground or in the air. \\
+from the x-y plane. Figure~\ref{fig:foot} shows the orientations of the x,y,z planes
+relative to a human foot in our calculation. Our implementation is a simplified version of
+the technique used in~\cite{fancy}. Figure~\ref{fig:step-detect} shows the basic shape of
+accelerometer data gathered from a subject's feet as he or she walks and the marked
+regions indicate the time during which his/her foot is either on the ground or in the air.
 \begin{figure}[h]
   \centering
-  \includegraphics[width=0.8\columnwidth]{figs/swing}
-  \caption{{\bf Step detection from accelerometer data}}
+  \subfigure[]{
+    \centering
+    \includegraphics[width=.95\columnwidth]{figs/swing}
     \label{fig:step-detect}
+  }
+  \subfigure[]{
+    \centering
+    \includegraphics[width=.7\columnwidth]{figs/one_foot}
+    \label{fig:foot}
+  }
+  \label{fig:swing}
+  \caption{{\bf Step detection graph and IMU references}}
 \end{figure}
+
 {\bf Knee Extension} A subject's knee extension was calculated using the hinge joint
-techniques explained by Yost~\cite{yost}. The equations are \utodo{Iljoo}{Add the
-equations you actually used for knee extension} Figure~\ref{fig:knees} shows the angles
-that are reported by the final calculation.
+techniques explained by Yost~\cite{yost}. The equations used to calculate the knee
+extension are shown on the right of Figure~\ref{fig:knees}. The picture on the left shows
+the angles that are reported by the final calculation and the orientations reported by
+each of the IMUs used in the calculation. Each quaternion from the IMUs is converted to a
+local coordinate using rotation matrix respectively. The angle is calculated by dot
+product of two quaternions projected to the x-axis in the IMU’s the local coordinate
+system.
 
@@
   \centering
-  \includegraphics[width=0.8\columnwidth]{figs/knees}
+  \includegraphics[width=0.99\columnwidth]{figs/knees2}
   \caption{{\bf Knee extension Calculation}}
@@
 \subsection{User Interface}
-\utodo{Alex}{Add all the details about the user interface that I missed!!}
-The user interface is aimed at a technical professional rather than a healthcare provider,
+The user interface is aimed at a technical professional rather that a healthcare provider,
 but it is built so that certain features are easy to access for non-technical personnel
@@
 reports whether or not the master node has successfully received data from a given slave
-within a recent timeframe. Figure~\ref{fig:ROS} shows the configuration interface and the
+within a recent time-frame. Figure~\ref{fig:ROS} shows the configuration interface and the
 health status indicator. In a healthcare environment, these tools would be used by
-technical personnell to configure the system software before it is used by a healthcare
+technical personnel to configure the system software before it is used by a healthcare
 provider.
@@
   \caption{{\bf ROS user interface}}
-  \label{fig:gui}
+  \label{fig:ROS}
 \end{figure}
@@
 straps held the slave devices securely in place and they did not restrict the subject's
-movements. 
+movements. Figure~\ref{fig:attached} shows the devices attached to a subject;
+Figure~\ref{fig:on_shoe} provides a close up of how the device connects to the patient's
+shoe.  
 %Except for that part where the ones on the thighs acted like tourniquets... but we won't
@@
     \includegraphics[width=.45\columnwidth]{figs/box}
-    \label{fig:box2}
   }
@@
     \includegraphics[width=.45\columnwidth]{figs/wLid}
-    \label{fig:wLid}
   }
+  \caption{{\bf Enclosure for sensor nodes}
   \label{fig:box}
-  \caption{{\bf Enclosure for sensor nodes}}
+  }
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \subfigure[]{
+    \centering
+    \includegraphics[width=.45\columnwidth]{figs/on_leg}
+    \label{fig:on_body}
+  }
+  \subfigure[]{
+    \centering
+    \includegraphics[width=.45\columnwidth]{figs/on_foot}
+    \label{fig:on_shoe}
+  }
+  \caption{{\bf Sensors attached to subject}
+  \label{fig:attached}
+  }
 \end{figure}
@@
   \caption{{\bf Final visualization layout}}
-  \label{fig:vis }
+  \label{fig:vis}
 \end{figure}
@@
 \subsection{Gait parameter graphs}
-In adition to visualizing a subject's movements, our system can generate graphs of a
+In addition to visualizing a subject's movements, our system can generate graphs of a
 subject's knee extension and stride length in real time. Figure~\ref{fig:extend} shows the
@@
   \caption{{\bf Real time knee extension reporting}}
-  \label{fig:extend }
+  \label{fig:extend}
 \end{figure}
@@
 in Figure~\ref{fig:falling}. Though our system relies primarily on the obvious change in
-the grph that a system operator can visually observe, an automatic mechanism based on a
+the graph that a system operator can visually observe, an automatic mechanism based on a
 set threshold for the angle between the nodes and the ground would suffice to warn
@@
 
-\balance
+\section*{Acknowledgements}
+We thank the Arjit Arora, Robotics Engineer, 4moms for assistance with design and
+fabrication of enclosures for sensor nodes. We would also like to show our gratitude to
+the Dr. Seungmoon Song and Dr. Hanbyul Joo, Robotics Institute, Carnegie Mellon University
+for sharing their pearls of wisdom with us during the course of this project. We are also
+immensely grateful to Prof. Rajkumar, Shunsuke Aoki, and Sandeep Dsouza for their
+support.
+
+%\balance
 \bibliographystyle{abbrv}
